### 2020-05-11 inst/app/www/on_load_placeholder.Rmd
working plotly version

---
title: ''
output: html_document
---

+ Please set the parameters and run the sample size planning calculation with the "Calculate sample size" button.
+ After running the calculation you can download the results in a general template text in word format.
+ After donwload, please complement the text where it is need with your reasoning behind each decision and include the text in your manuscript.