### 2020-06-28 inst/app/www/aipe_output.Rmd
working baseline version

---
title: "&nbsp;"
output:
  html_document: default
  word_document: default
params:
  n_min: NA
  n_max: NA
  output_n: NA
  input_power: NA
  output_confidence: NA
  band: NA
  delta: NA
---

```{r, include = FALSE}
library(knitr)
```

In order to calculate the sample size we choose the accuracy in parameter estimation (AIPE; REF) method. We choose the expected effect size to be `r expression (delta)` = `r params$delta`, because ... The desired power was `r params$input_power` as ... Because of limits in resources the minimum sample size of interest was `r params$n_min` and the maximum sample size of interest was `r params$n_max`. The resulting confidence of the sample size estimation method was `r params$output_confidence` with the estimated sample size of `r params$output_n`.