### 2020-07-03 inst/app/www/on_load_placeholder.Rmd

---
title: "&nbsp;"
output: html_document
---

+ Please set the parameters and run the sample size planning calculation with the "Calculate sample size" button.
+ After running the calculation you can download the results in a general template text in word format.
+ After download, please complement the text where it is need with your reasoning behind each decision and include the text in your manuscript.
+ Please wait for a calulation to finish before you attempt to download the output or open a new tab.