### 2021-03-19 inst/app/www/introduction.Rmd
updated UI and function documentation

---
title: ''
output: html_document
---
<center>
### How to use the app

This application provides 9 different methods to determine the sample size for a two group design. Moreover, the app provides help to create an accompanying report text for your sample size determination.

The figure below shows the 9 sample size estimation methods in the app and their relations. If you would like to know more about each method you can read our [tutorials manuscript](http://corndog.io/).

To get started with the sample size estimation choose a method under the __Methods__ tab.

<img src="www/flowchart.jpg" width="80%"/>

Please visit the [About](http://corndog.io/) tab if you would like to know more about the project or if you have any suggestions or issues with the app.

</center>