### 2021-03-30 README.Rmd
last update before submission

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SampleSizePlanner

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of SampleSizePlanner is to help researchers determine the sample size for two group designs. The present application and R package offers 9 different sample size planning methods. In addition, the web application allows users to create a justification report for their sample size plan.

## Available sample size planning methods
* Two One‐Sided Tests (TOST)
* Interval Equivalence Bayes factor
* Classical power analysis
* Power curve
* Bayes Factor Design Analysis (BFDA)
* Predetermined sample size with Bayes factor
* Accuracy In Parameter Estimation (AIPE)
* A-priori precision (APP)
* Region of Practical Equivalence (ROPE)

## Usage

SampleSizePlanner can be used either via the web app or via R.

### Using the web app

You can use the app at [https://martonbalazskovacs.shinyapps.io/SampleSizePlanner/](https://martonbalazskovacs.shinyapps.io/SampleSizePlanner/).

You can alternatively run the app locally on your own computer by following these instructions:

Install the development version (SampleSizePlanner is not available from CRAN) from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("marton-balazs-kovacs/SampleSizePlanner")
```

Running the app.

``` r
SampleSizePlanner::run_app()
```

### Using the package

You can use the R package to determine the sample size with any of the 9 methods that present in the web app. However, it is not possible currently to generate a justification report from the R package. Please see the tutorial paper or the web app for example sample size justifications for each method.

## Note

The tutorial paper that presents the web application and R package is not reviewed yet. More details are available in Kovacs, van Ravenzwaaij, Hoekstra & Aczel (2021).

## Code of Conduct

We are open to new ideas and feature requests.

Please note that the SampleSizePlanner project is released with a [Contributor Code of Conduct](https://marton-balazs-kovacs.github.io/SampleSizePlanner/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.