### 2021-03-30 inst/app/www/about.Rmd

---
title: "&nbsp;"
output: html_document
---

### About the project

The __SampleSizePlanner__ shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R. The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the chosen method. Therefore, a formatted template text containing the used parameters and the output of the calculations is available for download. This text can be complemented with the reasoning behind each decision.

### Contributors
- Marton Kovacs
- Don van Ravenzwaaij
- Rink Hoekstra
- Balazs Aczel

### Questions and comments

If you have any questions or comments please write to Marton Kovacs to the following email address: marton.balazs.kovacs@gmail.com

Or submit an issue on the apps Github page: [https://github.com/marton-balazs-kovacs/SampleSizePlanner](https://github.com/marton-balazs-kovacs/SampleSizePlanner)