### 2021-09-02 inst/app/www/about.Rmd
renaming, tidying, recalculating ssp results, debugging

---
title: "&nbsp;"
output: html_document
---

### About the project

The **SampleSizePlanner** shiny application enables researchers to calculate the sample size with 9 different methods in app and provides code to use an additional method from R. The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the chosen method. Therefore, a formatted template text containing the used parameters and the output of the calculations is available for download. This text can be complemented with the reasoning behind each decision.

### Contributors

-   Marton Kovacs
-   Don van Ravenzwaaij
-   Rink Hoekstra
-   Balazs Aczel

The design of the app was done by Zsuzsa Szekely [(@szekely_Zsuzsa)](<https://twitter.com/szekely_zsuzsa>).

### Questions and comments

If you have any questions or comments please write to Marton Kovacs to the following email address: [marton.balazs.kovacs@gmail.com](mailto:marton.balazs.kovacs@gmail.com)

Or submit an issue on the apps Github page: <https://github.com/marton-balazs-kovacs/SampleSizePlanner>