### 2024-11-27 inst/app/www/about.Rmd
Save local changes

---
title: "&nbsp;"
output: html_document
---

### About the project

The **SampleSizePlanner** shiny application enables researchers to calculate the sample size with 13 different methods in app and provides code to use an additional method from R. The aim of the app is to make the various sample size calculating methods available for a wide range of researchers and propagate the transparent reporting of the chosen method. Therefore, a formatted template text containing the used parameters and the output of the calculations is available for download. This text can be complemented with the reasoning behind each decision.

### Contributors

-   Marton Kovacs
-   Don van Ravenzwaaij
-   Rink Hoekstra
-   Balazs Aczel
-   Andrew Setiono
-   Sebastian Lortz

The design of the app was done by Zsuzsa Szekely [(@szekely_Zsuzsa)](<https://twitter.com/szekely_zsuzsa>).

### Questions and comments

If you have any questions or comments please write to Marton Kovacs to the following email address: [marton.balazs.kovacs@gmail.com](mailto:marton.balazs.kovacs@gmail.com)

Or submit an issue on the apps Github page: <https://github.com/marton-balazs-kovacs/SampleSizePlanner>